---
title: "Power Analysis"
author: "Ying"
date: "11/18/2019"
output:
  pdf_document:
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, include=FALSE}
library("tidyverse")
library("magrittr")
library("ggeffects")
library("broom")     # for elegant handling of model outputs
library("lme4")
library("lmerTest")
library("brms")

theme_set(theme_bw())
```

## Brief description 

We examine reaction times (RT) of a simple visual discrimination task that will be used as part of a larger study on duration thresholds. In this task, participants saw a Gabor patch that varied in 8 different contrasts (2.1, 3.6467, 6.3326, 10.9968, 19.0964, 33.1615, 57.5860, 100%). Each trial starts with a shrinking circle fixation. Then the stimulus is presented until the participants makes a key press. Participants have to discriminate the orientation of the stimulus (either tilted left or right). The size and the orientation of the gabor patch are held constant across trials and participants. 

The goal of this power analysis is to determine how many subjects will be required in order to reach 90% power in a paired t-test.

### Pilot data
Number of subjects = 2

Number of trials per subject = 1600: 2 (orientation) x 8 (contrasts) design with 100 trials/condition



### Variable descriptions:
*angle* = -30 (left tilt) or 30 (right tilt)

*RT* = reaction time in seconds

*response* = participant's response, 1 = left tilt, 2 = right tilt

*angle code* = recode angles -30 and 30 into 1 and 2

*accuracy* = whether the participant is correct, 1 = correct, 0 = incorrect


```{r load data, echo=FALSE}
d = read.csv("./static_data.csv")
```

## Scatter Plot

```{r descriptives}

d.summary = d %>%
  group_by(subjectID, contrasts) %>%
  summarise(meanRT = mean(RT)) 

ggplot(d.summary, aes(x = contrasts, y = meanRT, color = as.factor(subjectID))) + 
  geom_point() + 
  theme_classic() + labs(title = "Scatter plot of mean RT as function of contrasts") + 
  theme(plot.title = element_text(hjust=0.5, size = 15)) +
  labs(color = "Subjects")

```

