---
title: "BCS 511 Minilab 2 - Psychophysics: Sex-related differences in motion processing?"
author: "Ying Lin"
date: "11/1/2019"
output:
  pdf_document:
    latex_engine: xelatex
    number_sections: no
    toc: no
  word_document:
    toc: no
  html_document:
    df_print: paged
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, results = "markup",
  fig.align = "center", fig.width = 8)

options(width = 100, mc.cores = 4)

```

```{r libraries, include=FALSE}
library("tidyverse")
library("magrittr")
library("ggeffects")
library("broom")     # for elegant handling of model outputs
library("lme4")
library("lmerTest")
library("brms")

theme_set(theme_bw())
```

```{r functions, include=FALSE}
myCenter = function(x) { return(x - mean(x, na.rm = T)) }


prepVars = function(d) {
  d %<>%
    mutate_at(
      vars(
        one_of(
          c(
            "Contrast", 
            "LogDuration", 
            "EmphatizingQuotient", 
            "SystemizingQuotient", 
            "AutismQuotient", 
            "ADHD"))),
      .funs = funs(c = myCenter)
    ) %>%
    mutate_at(
      c("Sex", "VideoGaming"),
      .fun = factor
    )
  
  contrasts(d$Sex) = cbind("Female vs. Male" = c(-1,1))
  
  return(d)
}
```

```{r load data, include = FALSE, comment=NA, warning=FALSE}
load(file = "./data/Minilab2 data-10-30-19.RData")
```

# 1. Descriptive plots and analysis
## Visualization of duration thereshold and the other parameters (*Rmax*, *Slope*, and *C50*). Analyzed sex-related effects of all parameters using *t-test* 

We found that the effect of gender for all parameters were significant. 

```{r data formating, include=FALSE, comment=NA, warning=FALSE}
# omit participant 29 with NA vallues
d = na.omit(d)
d.all = d  #keeping a set of original data

# filter out motion task only
d %<>%
  filter(Task == "Motion")

# summary data by participant and contrasts
d.summary =
  d %>%
  group_by(Subject, Task, Contrast, Rmax, Slope, C50) %>%
  summarise(meanLogDuration = mean(LogDuration))

d.summary$Contrast = as.factor(d.summary$Contrast)

# separate duration summary with parameter summary
d.contrast_summary =
  d %>% 
  group_by(Subject, Contrast, Sex) %>%
  summarise(meanLogDuration = mean(LogDuration))

d.parameter_summary =
  d %>%
  group_by(Subject, Sex, Rmax, Slope, C50, Asymp, LogAsymp, LogC50, Contrast) %>%
  summarise()

d.parameter_summary$Sex = as.factor(d.parameter_summary$Sex)

```

```{r plots, comment=NA, warning=FALSE}

# # violin plot of mean of LogDuration
# ggplot(d.summary, aes(x = Contrast, y = meanLogDuration, color=Contrast)) + 
#   geom_violin() + geom_point() + theme_classic() + 
#   labs(title = "Violin plot of distribution of Duration Threshold by levels of Contrasts") + 
#   theme(plot.title = element_text(hjust=0.5, size = 20)) +
#   scale_x_discrete("Stimulus Contrasts") +
#   scale_y_continuous("Log Duration")

# histogram plot for parameters: Rmax, Slope, C50
ggplot(d.parameter_summary, aes(x = Contrast, fill= Sex)) + geom_histogram() +
  theme_classic() + labs(title = "Histogram distribution of Contrast") + 
  theme(plot.title = element_text(hjust=0.5, size = 20)) +
  scale_x_continuous("Contrast") +
  scale_y_continuous("Frequency")

ggplot(d.parameter_summary, aes(x = Rmax, fill= Sex)) + geom_histogram() +
  theme_classic() + labs(title = "Histogram distribution of Rmax") + 
  theme(plot.title = element_text(hjust=0.5, size = 20)) +
  scale_x_continuous("Rmax") +
  scale_y_continuous("Frequency")

ggplot(d.parameter_summary, aes(x = Slope, fill= Sex)) + geom_histogram() +
  theme_classic() + labs(title = "Histogram distribution of Slope") + 
  theme(plot.title = element_text(hjust=0.5, size = 20)) +
  scale_x_continuous("Slope") +
  scale_y_continuous("Frequency")

ggplot(d.parameter_summary, aes(x = C50, fill= Sex)) + geom_histogram() +
  theme_classic() + labs(title = "Histogram distribution of C50") + 
  theme(plot.title = element_text(hjust=0.5, size = 20)) +
  scale_x_continuous("C50") +
  scale_y_continuous("Frequency")
```

## Analysis: comparing effects of gender on LogDuration, Rmax, Slope, and C50

```{r t-tests, comment=NA, warning=FALSE}

# lm and t-tests 
# tidy(glm(data = d.contrast_summary, meanLogDuration~Contrast + as.factor(Sex))) # incorrect 
t.test(data = d.parameter_summary, Contrast~as.factor(Sex)) # not sig
t.test(data = d.parameter_summary, Rmax~as.factor(Sex))     # sig
t.test(data = d.parameter_summary, Slope~as.factor(Sex))    # sig
t.test(data = d.parameter_summary, C50~as.factor(Sex))      # sig

```

# 2. Using alternative measures of parameter Rmax and C50 (*Asymp* and *LogAsymp* instead of *Rmax*; and *LogC50* instead of *C50*)

```{r t-tests2, comment=NA, warning=FALSE}
# t-tests
t.test(data = d.parameter_summary, Asymp~as.factor(Sex))    # sig
t.test(data = d.parameter_summary, LogAsymp~as.factor(Sex)) # no sig
t.test(data = d.parameter_summary, LogC50~as.factor(Sex))   # sig
  
# Analysis of paramater after filtering out +/- 3SD of mean 

# Rmax outliers
d.out_rmax = d.parameter_summary %>%
  mutate(rmax_in = as.factor( abs(d.parameter_summary$Rmax) < 3*sd(d.parameter_summary$Rmax) + mean(d.parameter_summary$Rmax))) %>%
  filter(rmax_in) #error

t.test(data = d.out_rmax, Rmax~as.factor(Sex))  #error

# Slope outliers
d.out_slope = d.parameter_summary %>%
  filter(Slope < 3*sd(Slope))

t.test(data = d.out_slope, Slope~as.factor(Sex))  #error

# C50 outliers
d.out_c50 = d.parameter_summary %>%
  filter(C50 < 3*sd(C50))

t.test(data = d.out_c50, C50~as.factor(Sex))  #error

```

# 3. Anova - testing whether sex-related effects differ significantly between the four parameters

```{r anova, comment=NA, warning=False}

a <- anova(Rmax ~ Slope + C50, d.parameter_summary)
tidy(a)

```


# References

 * Wheelwright, S., Baron-Cohen, S., Goldenfeld, N., Delaney, J., Fine, D., Smith, R., ... & Wakabayashi, A., 2006). Predicting autism spectrum quotient (AQ) from the systemizing quotient-revised (SQ-R) and empathy quotient (EQ). Brain research, 1079(1), 47-56

# Session info
```{r session_info, echo=FALSE, results='markup'}
devtools::session_info()
```
