---
title: "BCS 511 Minilab 2 - Psychophysics: Sex-related differences in motion processing?"
author: "Ying Lin"
date: "11/1/2019"
output:
  pdf_document:
    latex_engine: xelatex
    number_sections: no
    toc: no
  word_document:
    toc: no
  html_document:
    df_print: paged
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, results = "markup",
  fig.align = "center", fig.width = 8)

options(width = 100, mc.cores = 4)

```

```{r libraries, include=FALSE}
library("tidyverse")
library("magrittr")
library("ggeffects")
library("broom")     # for elegant handling of model outputs
library("lme4")
library("lmerTest")
library("brms")

theme_set(theme_bw())
```

```{r functions, include=FALSE}
myCenter = function(x) { return(x - mean(x, na.rm = T)) }


prepVars = function(d) {
  d %<>%
    mutate_at(
      vars(
        one_of(
          c(
            "Contrast", 
            "LogDuration", 
            "EmphatizingQuotient", 
            "SystemizingQuotient", 
            "AutismQuotient", 
            "ADHD"))),
      .funs = funs(c = myCenter)
    ) %>%
    mutate_at(
      c("Sex", "VideoGaming"),
      .fun = factor
    )
  
  contrasts(d$Sex) = cbind("Female vs. Male" = c(-1,1))
  
  return(d)
}
```

```{r load data, include = FALSE, comment=NA, warning=FALSE}
load(file = "./data/Minilab2 data-10-30-19.RData")
```

# 1. Descriptive plots and analysis
## Visualization of parameters (*Rmax*, *Slope*, and *C50*).

```{r data formating, include=FALSE, comment=NA, warning=FALSE}
# omit NA vallues
d = na.omit(d)
d.all = d  #keeping a set of original data

# filter out motion task only
d %<>%
  filter(Task == "Motion")

# summary data 
d.parameter_summary =
  d %>%
  group_by(Subject, Sex, Rmax, Slope, C50, Asymp, LogAsymp, LogC50, VideoGaming) %>%
  summarise()

```

```{r plots, comment=NA, warning=FALSE}

ggplot(d.parameter_summary, aes(x = Rmax, fill= as.factor(Sex))) + 
  geom_histogram(alpha = 0.8, position = position_dodge()) +
  theme_classic() + labs(title = "Histogram distribution of Rmax") + 
  theme(plot.title = element_text(hjust=0.5, size = 20)) +
  scale_x_continuous("Rmax") +
  scale_y_continuous("Frequency")

ggplot(d.parameter_summary, aes(x = Slope, fill= as.factor(Sex))) + geom_histogram(alpha = 0.8, position = position_dodge()) +
  theme_classic() + labs(title = "Histogram distribution of Slope") + 
  theme(plot.title = element_text(hjust=0.5, size = 20)) +
  scale_x_continuous("Slope") +
  scale_y_continuous("Frequency")

ggplot(d.parameter_summary, aes(x = C50, fill= as.factor(Sex))) + geom_histogram(alpha = 0.8, position = position_dodge()) +
  theme_classic() + labs(title = "Histogram distribution of C50") + 
  theme(plot.title = element_text(hjust=0.5, size = 20)) +
  scale_x_continuous("C50") +
  scale_y_continuous("Frequency")
```

## Analysis: comparing effects of gender on Rmax, Slope, and C50
Using a linear regression model to access if gender (0 = male, 1 = female) has a significant effect on Rmax, Slope and C50.

*Rmax*: We found a significant difference of Rmax ($\hat\beta$ = -0.02, *t* = 0.005, *p* < .001) between men and women. Suggesting that women converage against higher minimal duration threshold in the motion processing task. 
*Slope*: We did not find a significant difference of Slopw ($\hat\beta$ = 0.72, *t* = 0.70, *p* > .05).
*C50*: Lastly, we also found a significant difference of C50 ($\hat\beta$ = -2.09, *t* = -2.41, *p* < 0.05), suggesting that women shows a lower C50 value than men. 

```{r lm, comment=NA, warning=FALSE}

# Rmax
rmax_model = lm(data = d.parameter_summary, Rmax~as.factor(Sex))
tidy(rmax_model)

# Slope
slope_model = lm(data = d.parameter_summary, Slope~as.factor(Sex)) 
tidy(slope_model)

# C50; depends on excluding outliers or not 
C50_model = lm(data = d.parameter_summary, C50~as.factor(Sex)) 
tidy(C50_model)

d.parameter_summary %>%
  ungroup() %>%
  mutate(s.C50 = as.numeric(scale(C50)))
```

# 2. Using alternative measures of parameter Rmax and C50 (*Asymp* and *LogAsymp* instead of *Rmax*; and *LogC50* instead of *C50*)

Using alternative measures of parameters we found a significant sex difference between men and women for 
Log of C50 (*t* = -4.60 ,*p* < 0.05) and Asymp (*t* = -4.06, *p* < 0.05) but not for Log of Asymp (*t* = 0.40, *p* > 0.05). This suggests that different transformations of a parameter does influence the significance of a finding. 

We also examined outliers of +/- 3SD from the mean for Rmax, Slope, and C50. The significance of the linear model still holds for Rmax and C50 after removal of outliers.

```{r t-tests2, comment=NA, warning=FALSE}

# t-tests
t.test(data = d.parameter_summary, Asymp~as.factor(Sex))    # 1/rmax 
t.test(data = d.parameter_summary, LogAsymp~as.factor(Sex)) # log 1/rmax - duration is usually analyze by log duration (aka, -log rmax)
t.test(data = d.parameter_summary, LogC50~as.factor(Sex))   # log C50
  
# Analysis of paramater after filtering out +/- 3SD of mean 

# no outliers for Rmax (within 3SD); transformation will not be the same 
d.out_rmax = d.parameter_summary %>%
  ungroup() %>%
  mutate(rmax_in = abs(scale(Rmax)) < 3) %>% 
  filter(rmax_in == 1)
  # mutate(rmax_in = as.integer(as.logical(abs(d.parameter_summary$Rmax - mean(d.parameter_summary$Rmax)) 
                                         # < 3*sd(d.parameter_summary$Rmax)))) %>%
  
  tidy(lm(data = d.out_rmax, Rmax~as.factor(Sex)))
  
  
d.out_slope = d.parameter_summary %>%
  ungroup() %>%
  mutate(slope_in = abs(scale(Slope)) < 3) %>% 
  filter(slope_in == 1)

  tidy(lm(data = d.out_slope, Slope~as.factor(Sex)))
  
  
d.out_c50 = d.parameter_summary %>%
  ungroup() %>%
  mutate(c50_in = abs(scale(C50)) < 3) %>% 
  filter(c50_in == 1)

  tidy(lm(data = d.out_slope, C50~as.factor(Sex)))

```

# 3. Anova - testing whether sex-related effects differ significantly between the four parameters

```{r anova, comment=NA, warning=FALSE}
# only if the interaction is significant that sex effects of two measures are different  
anova(rmax_model, slope_model, test = "Chisq")
anova(rmax_model, C50_model, test = "Chisq")
anova(slope_model, C50_model, test = "Chisq")

```

# 5. Individual differences variables

```{r individual differences, comment=NA, warning = FALSE}
# DV ~(Sex + ADHD...) * measure --- how does individuals differences affect Rmax
# if sex becames non-significant, possibly that the effect is from (ex.)ADHD that is mediated by sex

# histogram plot for video game
ggplot(d.parameter_summary, aes(x = Rmax, fill= as.factor(VideoGaming))) + geom_histogram(alpha = 0.8, position = position_dodge()) +
  theme_classic() + labs(title = "Histogram distribution of Rmax") + 
  theme(plot.title = element_text(hjust=0.5, size = 20)) +
  scale_x_continuous("Rmax") +
  scale_y_continuous("Frequency")

# include other variables
```

# References

 * Wheelwright, S., Baron-Cohen, S., Goldenfeld, N., Delaney, J., Fine, D., Smith, R., ... & Wakabayashi, A., 2006). Predicting autism spectrum quotient (AQ) from the systemizing quotient-revised (SQ-R) and empathy quotient (EQ). Brain research, 1079(1), 47-56

# Session info
```{r session_info, echo=FALSE, results='markup'}
devtools::session_info()
```
